---
title: "finance tree"
author: "surecalois"
date: "1/28/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# simulating how interest rate (forward rate) change

simulating one possible rate changes, at each period, the rate might go either up or down. the amount of up or down is controled by paramter `d`.

```{r}

simple.rate = function(n,d = 0.02){
  ch = sample(c(1,-1),n,replace = T) #equal chance going up and down
  r.m = c(1,exp(d*cumsum(ch)))
  return(r.m)
}

r.m = 0.06*simple.rate(30)
plot(r.m)
lines(r.m)

prob.rate = function(n,d = 0.02,p = 0.5){
  h = c(-1,1)
  x = runif(n) > (1-p)
  ch = h[x+1]
  #ch = sample(c(1,-1),n,replace = T) #equal chance going up and down
  r.m = c(1,exp(d*cumsum(ch)))
  return(r.m)
}

r.m = 0.06*prob.rate(30,p = 0.56)
plot(r.m)
lines(r.m)

```


```{r}

show_rate = function(r.m){
  r.m$t = 0:(length(r.m[[1]])-1)
  gr.m = gather(r.m, key = "key",value = "rate",-t)
  p = ggplot(data = gr.m) + geom_path(aes(x = t,y = rate, color = key))+
    scale_y_continuous(trans='log10')+
    theme(legend.position = "none")
  print(p)
}

term = 30
trial = 1:200
names(trial) = paste0("r",trial)
r.m = map_dfc(trial,~0.06*prob.rate(term,0.05,0.55))
show_rate(r.m)


```

```{r}
var_loan = function(coupon,term,rate = rate, face = 100){
  s = seq(0,term)
  pvs = s
  pvs[term+1] = face #after coupon payment
  for(i in term:1){
    pvs[i] = (coupon + pvs[i+1])/(1+rate[i])
  }
  names(pvs) = s
  return(rev(pvs))
}

trial = 1:1000
names(trial) = paste0("r",trial)
term = 30
r.m = map(trial,~0.08*prob.rate(term,0.16,0.5))

a = map_dfr(r.m,~var_loan(9,term,rate = .x))
y = a$`0`
my = mean(y)
sy = sd(y)
hist(y,breaks = 30)
abline(v = my,col = "red")
abline(v=c(my-sy,my+sy),col = "red",lt = 4)
```


simulating the whole picture of interest rate change

```{r}
full.rate = function(term,d = 0.16){
  
  ff = function(k){
    a = seq(0,k)
    ch = a - rev(a)
    return(exp(d*ch))
  }
  
  s = 0:term
  names(s) = s
  rates = map(s,ff)
  return(rates)
}

```

some helper functions
```{r}
tidylist = function(r.m){
  x = names(r.m)
  xs = map(x,~rep(.x,length(r.m[[.x]])))
  pts.x = flatten_chr(xs)
  pts.x = factor(pts.x,levels = x)
  pts.y = flatten_dbl(r.m)
  pts = data.frame(key = pts.x, value = pts.y)
  return(pts)
}

make_fr.tree = function(r.m){
 term = length(r.m)-1
 pts = tidylist(r.m)
 pts$key = as.numeric(pts$key)
 names(pts)=c("x","y")
 ind = seq(1,term+1)
 edge1 = cumsum(ind) #depends on the full.rate put up first or down first
 edge2 = edge1 - ind + 1
 names(ind) = ind
 maz = function(x,th){x[x>=th]}
 gg1 = map(ind,~maz(edge1-.x+1,edge2))
 miz = function(x,th){x[x<=th]}
 gg2 = map(ind,~miz(edge2+.x-1,edge1))
 
 # plot(pts,log = 'y')
 # walk(gg2,~lines(pts[.x,],col = "blue"))
 # walk(gg1,~lines(pts[.x,],col = "red"))

 gg1.d= tidylist(gg1)
 gg2.d= tidylist(gg2)
 
 pts$gg1 = 0
 pts[gg1.d$value,'gg1'] = gg1.d$key
 pts$gg1 = factor(pts$gg1,levels = levels(gg1.d$key))
 pts$gg2 = 0
 pts[gg2.d$value,'gg2'] = gg2.d$key
 pts$gg2 = factor(pts$gg2,levels = levels(gg2.d$key))
 
 return(pts)
}

```

```{r}
r.m = full.rate(30)
r.m = map(r.m,~0.06*.x)

pts = make_fr.tree(r.m)

p = ggplot(data = pts)+
  geom_point(aes(x = x,y = y))+
  geom_path(aes(x = x,y =y,group = gg1),color = "lightblue")+
  geom_path(aes(x = x,y =y,group = gg2),color = "pink")+
  scale_y_continuous(trans = 'log10')
print(p)


```


to show the probability of the scenarios
```{r}

bf.tree = function(r.m,p){
  szs = as.numeric(names(r.m))
  names(szs) = names(r.m)
  ps = map(szs,~dbinom(0:.x,.x,p))
  return(ps)
}

ps = bf.tree(r.m,0.55)
ps.d = tidylist(ps)
names(ps.d) = c("x","p")
pts$p = ps.d$p #left_join won't work because x is not unique.

p = ggplot(data = pts)+
  geom_point(aes(x = x,y = y,color = p))+
  scale_color_continuous(type = "viridis")+
  scale_y_continuous(trans = 'log10')
print(p)

pts = pts %>% group_by(x) %>% mutate(np = p/max(p))

p = ggplot(data = pts)+
  geom_point(aes(x = x,y = y,color = np))+
  scale_color_continuous(type = "viridis")+
  scale_y_continuous(trans = 'log10')
print(p)
```

filter out the less likely scenarios
```{r}
p = ggplot(data = pts %>% filter(p > 0.05))+
  geom_point(aes(x = x,y = y,color = np))+
  scale_color_continuous(type = "viridis")+
  scale_y_continuous(trans = 'log10',limits=c(1e-3, 1e1))
print(p)

```

# how to value a cashflow

two simple types of fixed rate cash flow

 - coupon bond
 - mortgage

```{r}
simple_bond = function(coupon,term,rate = 0.08, face = 100){
  s = seq(0,term)
  pvs = s
  pvs[term+1] = face #after coupon payment
  for(i in term:1){
    pvs[i] = (coupon + pvs[i+1])/(1+rate)
  }
  names(pvs) = s
  return(pvs)
}


mortgage = function(term,rate,amount=100){
  s = (1+rate)^(1:term)
  pay = amount/sum(1/s)
  balance = rep(0,term+1)
  for(i in term:1){
    balance[i] = (pay + balance[i+1])/(1+rate)
  }
  names(balance) = seq(0,term)
  pays = c(-amount,rep(pay,term))
  names(pays) = seq(0,term)
  return(list(payment = pays, balance = balance))
}
```

```{r}
z = mortgage(30,0.08,100)#30 year, 8%, 100
mg.flow = z$payment
print(z$balance)
barplot(mg.flow)

s = simple_bond(9,30,0.08,100)#9x30 year, 8%, 100
sl.flow = c(-s[1],rep(9,30-1),109)
barplot(sl.flow)
```

# some trees
- interest rate tree
  the middle nodes are fused together. the number of leaves grow linear with levels
  the number of paths back to the root are those "binomial number".
- interest rate path tree
  each nodes has only one path back to the root.


# the bookies and hedging

```{r}
binomial.tree = function(end.list,p){
  term = length(end.list)-1
  s = term:0
  names(s) = s
  pvs = list()
  pvs[[as.character(term)]] = end.list
  for(k in s[-1]){
    a = pvs[[as.character(k+1)]]
    b = p*a[1:(k+1)]+(1-p)*a[2:(k+2)]
    pvs[[as.character(k)]] = b
  }
  return(rev(pvs))
}

pl = c(rep(100,4),rep(-100,4))
bt = binomial.tree(pl,0.6)
bt.d = make_fr.tree(bt)

p = ggplot(data = bt.d)+
  geom_path(aes(x = x,y =y,group =gg1),color = "lightblue")+
  geom_path(aes(x = x,y =y,group =gg2),color = "pink")+
  geom_point(aes(x = x,y = y))
print(p)

plot(tidylist(bt))

a =  sum(dbinom(4:7,7,0.6))*100
b = 100-a
kip = a-b #case 1:1
k = (100 -kip)/b #k = (100+kip)/a should give the same
pl = c(rep(-b,4),rep(a,4))*k

bt2 = binomial.tree(pl,0.6)
bt2.d = make_fr.tree(bt2)

p = ggplot(data = bt2.d)+
  geom_path(aes(x = x,y =y,group =gg1),color = "lightblue")+
  geom_path(aes(x = x,y =y,group =gg2),color = "pink")+
  geom_point(aes(x = x,y = y))
print(p)


bt.d$g = 0
bt2.d$g = 1
bt3.d = bt.d
bt3.d$g = 2
bt3.d$y = bt.d$y+bt2.d$y
data = rbind(bt.d,bt2.d,bt3.d)
p = ggplot()+
  geom_path(data = data,aes(x = x,y =y,group =gg1),color = "blue",linetype = "11")+
  geom_path(data = data,aes(x = x,y =y,group =gg2),color = "red",linetype = "88")+
  geom_point(data = data,aes(x = x,y = y))+facet_grid(g~.)
print(p)

```


the two trees are identical, only the second one is flipped. when the `pl` at the end of the tree has the same geometry will result in the same tree geometry.

```{r}

pl =runif(8,min = -100,max = 100)
pl = sort(pl)
pl2 = 42 - pl
bt1 = binomial.tree(pl,0.6)
bt2 = binomial.tree(pl2,0.6)

bt1.d = make_fr.tree(bt1)
bt2.d = make_fr.tree(bt2)

bt1.d$g = 0
bt2.d$g = 1
bt3.d = bt1.d
bt3.d$g = 2
bt3.d$y = bt1.d$y+bt2.d$y
data = rbind(bt1.d,bt2.d,bt3.d)
p = ggplot()+
  geom_path(data = data,aes(x = x,y =y,group =gg1),color = "blue",linetype = "11")+
  geom_path(data = data,aes(x = x,y =y,group =gg2),color = "red",linetype = "88")+
  geom_point(data = data,aes(x = x,y = y))+facet_grid(g~.)
print(p)
```

# imperfect hedging
```{r}

pl =runif(8,min = -100,max = 100)
pl = sort(pl)
pl2 = 42 - pl+rnorm(8,sd = 5)
bt1 = binomial.tree(pl,0.6)
bt2 = binomial.tree(pl2,0.6)

bt1.d = make_fr.tree(bt1)
bt2.d = make_fr.tree(bt2)

bt1.d$g = 0
bt2.d$g = 1
bt3.d = bt1.d
bt3.d$g = 2
bt3.d$y = bt1.d$y+bt2.d$y
data = rbind(bt1.d,bt2.d,bt3.d)
p = ggplot()+
  geom_path(data = data,aes(x = x,y =y,group =gg1),color = "blue",linetype = "11")+
  geom_path(data = data,aes(x = x,y =y,group =gg2),color = "red",linetype = "88")+
  geom_point(data = data,aes(x = x,y = y))+facet_grid(g~.)
print(p)
```
